---
title: "Chapter 1"
author: Logan | Nick
output: github_document
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.align="center", echo = FALSE)
```

```{r message=FALSE}
if (!require(mosaic)) install.packages("mosaic", repos = 'https://cloud.r-project.org')
library(mosaic) # load the package mosaic to use its functions
```

# 1.1 Populations, Samples, and Processes

Statistics provides methods for organizing and summarizing data and for drawing conclusions from that data

* Def

Data : a collection of facts

* Def

Population : A well defined collection of objects for which we wish to obtain info

* Def

Census : When desired info is obtained from every member of the population

 - problems : Time, money, practical

* Def

Sample : A subset of the population

1. 

You want the home price in Edwardsville

 - Fewer well trained appraisers gives better results than many poorly trained 

2. Tree Age Study

Testing is destructive, so a sample is better

* Def

variable : any characteristic whose clue may differ from one subject to another.

 - denote with low letters

**Note** 
    
    - Don't say $McDonald's = 10$
      
    - Do say $x =$ the length of the tibia bone in 10 year old boys.

* Def

univariate data : result from making observations of 1 variable

 - these variable can be qualitative / quantitative

* Def

Bivariate data : when observations are made on each of 2 variables for each individual 

 - (weight.mpg) of cars

* Def

Multivariate data : observations made on many variables

 - patient data


* Ex

Labor force, sample 60,000, find population  + sample

 - population = labor force, sample size = 60,000 households

## Branches of Stats
1. Descriptive Stats : data are collected and you wish to summarize and describe features of the data (graphs, numerical summaries)

2. Inferential stats : data is collected from a sample and used to draw a conclusion about the population

    - confidence intervals, hypothesis test, prediction, etc...

## Types of sampling

- Simple random sampling : random choice / draw of the hat sampling

- Systematic sampling : selecting every $k^{th}$ member of the population

- Cluster sampling : divide population into groups, then select some of these groups @ random

- Stratified sampling : divide population into groups. Find subgroups of groups (strata) and then draw random sample in strata

- Convenience sampling : sampling in the most convenient way 
  - best to avoid , but a good starter

## Notate

### sample size : $n$

 - For a dataset with $n$ observations on some variable $x$, the individual observations will be denoted as $x_1, x_2, \ldots, x_n$.

# 1.2 

## Stem and leaf plots 

* Ex

$(54, 59, 35, 41, 46, 25, 47, 60, 54, 46, 49, 46, 41, 34, 22)$

During these problems it helps to first orgranize the numebrs in the list first
\begin{align*}
    2 &\mid 2, 5 \\
    3 &\mid 4, 5 \\
    4 &\mid 1, 1, 6, 6, 6, 7, 9 \\
    5 &\mid 4, 4, 9 \\
    6 &\mid 0
\end{align*}

## Dot plots
```{r}
mlb16.data <- read.csv("https://raw.githubusercontent.com/jpailden/rstatlab/master/data/MLB-TeamBatting-S16.csv")
dotPlot(~BatAge, data = mlb16.data)
```

## Histograms
\color{red} left blank for nick to fill in

\color{black}

## Skewed (Right and left)
\color{red} add a dataset to show?

\color{black}


## Bell
\color{red} add a dataset to show?

\color{black}

## Flat uniform
\color{red} add a dataset to show?

\color{black}

## nonsymmetric
\color{red} add a dataset to show?

\color{black}

## bimodal symmetric
\color{red} add a dataset to show?

\color{black}

# 1.3

* Def

mean : numerical value of average

Notate 

### Sample mean : $\overline{x}$
\[
\overline{x} = \frac{x_1 + x_2 + \ldots + x_n}{n} = \frac{\sum\limits_{i=1}^{n} x_i}{n}
\] 

Notate

### Population mean : $\mu$

- avg of all values in the entire pop. 

* Ex

$2, 2, 5, 3, 8, 9, 2, 3, 1$
 \[
\overline{x} = \frac{\sum\limits_{i=1}^{10} x_i}{10} = 3.6
\] 

The mean is inappropriate in some cases b/c of outliers.

- this makes the mean a **nonresistant measure**  

* Def

Median : middle value /avg of 2 middle values when sorted

Notate

Median : $\widetilde{x}$ 

- if $n = odd$, median is at $\dfrac{n + 1}{2}$ 

- if $n = even$ median are b/n $\dfrac{n}{2}$ & $\dfrac{n + 1}{2}$ 

Notate

Population Mean : $\widetilde{\mu}$

# 1.4 Measures of Variability

One way to describe a distribution is by using the standard deviation

## Quartiles

-  $Q_1$ - lower quartile separates bottom $25\%$
-  $Q_2$ - median middle $50\%$
-  $Q_3$ - upper quartile separates upper $25\%$

* Ex

2, 2, 5, 1, 3, 8, 9, 2, 31

SORT
\begin{align*}
    1, 1, 2, 2, 2 \\
    3, 3, 5, 8, 9
\end{align*}
$\widetilde{x} = 2 + 3 = 2.5$

## Five number summary

- Find min, $Q_1$, median, $Q_3$, max

Note : If median is found in list, use it in both top half and lower half.

Ex : 2 2 5 1 3 8 9 2 3 1 100

$\overline{x} = \frac{36 + 100}{11} \approx 12.36$ 

Sort to find median. $\widetilde{x} = 3$.

## Mean vs. Median

- median is the equal parts point

- mean is the balance point

Notate

### Trimmed mean : $\overline{x}_{tr}$
 
- compromise b/n the mean & median

- to find it, remove top & bottom $10\%$, then calculate the mean
