---
title: "chapter 1"
author: "logan | nick"
output: github_document
fontsize: 12pt
---

```{r setup, include=false}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.align="center", echo = false)
```

```{r message=false}
if (!require(mosaic)) install.packages("mosaic", repos = 'https://cloud.r-project.org')
library(mosaic) # load the package mosaic to use its functions
```

# 1.1 populations, samples, and processes

statistics provides methods for organizing and summarizing data and for drawing conclusions from that data

* def

data : a collection of facts

* def

population : a well defined collection of objects for which we wish to obtain info

* def

census : when desired info is obtained from every member of the population

 - problems : time, money, practical

* def

sample : a subset of the population

1. 

you want the home price in edwardsville

 - fewer well trained appraisers gives better results than many poorly trained 

2. tree age study

testing is destructive, so a sample is better

* def

variable : any characteristic whose clue may differ from one subject to another.

 - denote with low letters

**note** 
    
    - don't say $mcdonald's = 10$
      
    - do say $x =$ the length of the tibia bone in 10 year old boys.

* def

univariate data : result from making observations of 1 variable

 - these variable can be qualitative / quantitative

* def

bivariate data : when observations are made on each of 2 variables for each individual 

 - (weight.mpg) of cars

* def

multivariate data : observations made on many variables

 - patient data


* ex

labor force, sample 60,000, find population  + sample

 - population = labor force, sample size = 60,000 households

## branches of stats
1. descriptive stats : data are collected and you wish to summarize and describe features of the data (graphs, numerical summaries)

2. inferential stats : data is collected from a sample and used to draw a conclusion about the population

    - confidence intervals, hypothesis test, prediction, etc...

## types of sampling

- simple random sampling : random choice / draw of the hat sampling

- systematic sampling : selecting every $k^{th}$ member of the population

- cluster sampling : divide population into groups, then select some of these groups @ random

- stratified sampling : divide population into groups. find subgroups of groups (strata) and then draw random sample in strata

- convenience sampling : sampling in the most convenient way 
  - best to avoid , but a good starter

## notate

### sample size : $n$

 - for a dataset with $n$ observations on some variable $x$, the individual observations will be denoted as $x_1, x_2, \ldots, x_n$.

# 1.2 

## stem and leaf plots 

* ex

$(54, 59, 35, 41, 46, 25, 47, 60, 54, 46, 49, 46, 41, 34, 22)$

during these problems it helps to first orgranize the numebrs in the list first
\begin{align*}
    2 &\mid 2, 5 \\
    3 &\mid 4, 5 \\
    4 &\mid 1, 1, 6, 6, 6, 7, 9 \\
    5 &\mid 4, 4, 9 \\
    6 &\mid 0
\end{align*}

## dot plots
```{r}
mlb16.data <- read.csv("https://raw.githubusercontent.com/jpailden/rstatlab/master/data/mlb-teambatting-s16.csv")
dotplot(~batage, data = mlb16.data)
```

## histograms
\color{red} left blank for nick to fill in

\color{black}

## skewed (right and left)
\color{red} add a dataset to show?

\color{black}


## bell
\color{red} add a dataset to show?

\color{black}

## flat uniform
\color{red} add a dataset to show?

\color{black}

## nonsymmetric
\color{red} add a dataset to show?

\color{black}

## bimodal symmetric
\color{red} add a dataset to show?

\color{black}

# 1.3

* def

mean : numerical value of average

notate 

### sample mean : $\overline{x}$
\[
\overline{x} = \frac{x_1 + x_2 + \ldots + x_n}{n} = \frac{\sum\limits_{i=1}^{n} x_i}{n}
\] 

notate

### population mean : $\mu$

- avg of all values in the entire pop. 

* ex

$2, 2, 5, 3, 8, 9, 2, 3, 1$
 \[
\overline{x} = \frac{\sum\limits_{i=1}^{10} x_i}{10} = 3.6
\] 

the mean is inappropriate in some cases b/c of outliers.

- this makes the mean a **nonresistant measure**  

* def

median : middle value /avg of 2 middle values when sorted

notate

median : $\widetilde{x}$ 

- if $n = odd$, median is at $\dfrac{n + 1}{2}$ 

- if $n = even$ median are b/n $\dfrac{n}{2}$ & $\dfrac{n + 1}{2}$ 

notate

population mean : $\widetilde{\mu}$

# 1.4 measures of variability

one way to describe a distribution is by using the standard deviation

## quartiles

-  $q_1$ - lower quartile separates bottom $25\%$
-  $q_2$ - median middle $50\%$
-  $q_3$ - upper quartile separates upper $25\%$

* ex

2, 2, 5, 1, 3, 8, 9, 2, 31

sort
\begin{align*}
    1, 1, 2, 2, 2 \\
    3, 3, 5, 8, 9
\end{align*}
$\widetilde{x} = 2 + 3 = 2.5$

## five number summary

- find min, $q_1$, median, $q_3$, max

note : if median is found in list, use it in both top half and lower half.

ex : 2 2 5 1 3 8 9 2 3 1 100

$\overline{x} = \frac{36 + 100}{11} \approx 12.36$ 

sort to find median. $\widetilde{x} = 3$.

## mean vs. median

- median is the equal parts point

- mean is the balance point

notate

### trimmed mean : $\overline{x}_{tr}$
 
- compromise b/n the mean & median

- to find it, remove top & bottom $10\%$, then calculate the mean

## categorical data

- the natural way to numerically summarize categorical data is by finding the \underline{proportion} of successes and failures

notate

### sample proportions : $\hat{p} = \dfrac{\text{\# of successes}}{n}$

notate

population proportions : $p =$ # of successes in the population

reporting a center of measure gives only partial info

sets mayh have similar means but differ in other ways

a simple way to give more detail is to give the \underline{range}

- def

range : max - min

deviations from the mean

- a dev from the mean is the absolute diference (distance) b.n an observation and the mean
 \[
x_1 - \overline{x}, x_2 - \overline{x}, \ldots, x_n - \overline{x}
\] 
note 
\[
\sum\limits_{i=1}^{n} (x_i - \overline{x}) = 0
\] 
\color{red} proof (omitted to catch up)

\color{black}

- def

standard deviation = measure of how much an observation is expected to be from the mean

notate

### population std. dev = $\sigma$

notate

### sample std. dev =
\[
s = \sqrt{\frac{\sum\limits_{i=1}^{n} (x_i - \overline{x})^2}{n - 1}} 
\] 
$\sigma$ is interpreted as size of typical deviation from $\mu$ w/ entire pop. of x-values

s has same units as data

note

s is not resistant (strongly affected by outliers / skew b.c of $\overline{x}$ )

$s \ge 0$ 

- def

### variance = $std.dev^2$ 

### pop variance = $\sigma^2$ 

### sample variance = $s^2$

