---
title: "Chapter 2"
author: "Logan | Nick"
output: github_document
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.align="center", echo = FALSE)
```

```{r message=FALSE}
if (!require(mosaic)) install.packages("mosaic", repos = 'https://cloud.r-project.org')
library(mosaic) # load the package mosaic to use its functions
```

##### Def

Experiment : any activity / situation w/ uncertainty about which $x \ge 2$ outcomes are possible 

- coin toss, roll a die, draw a card

##### Def

### Sample Space

- collection of all possible outcomes of a chance experiment 

Notate - s or $\mathcal{U}$

- Toss a coin : heads or tails

##### Def

#### Event : 

- any collection of outcomes from a sample space of a chance experiment

Notate

CAP letters : $A, B, C, \ldots$ 

##### Def

Simple Event : event that consists of \underline{one} outcome

Compound Event : event that consists of \underline{more than one} outcome

##### Ex

Tennis : A tennis shop carries 5 brands of rackets (Head, Prince, Sazenger, Wimbledon, Wilson). Each racket comes in midsize / oversize

a. sample space

*insert diagram here*

b. Let A be the event an oversized racket is piurchased
\[
A = \{HO, PO, SO, WimO, WilO\}
\] 

c. Let B be the event the name brand starts w/ a W
\[
 B = \{ WimM, WimO, WilM, WilO \}
\] 

### Forming New Sets

 Let A and B be any 2 events

##### Def

 Complement of A : 

 - all outcomes in S, not in A

##### Notate

$A^{\prime}, \overline{A}, A^c$

##### Notate

union - A or B - inclusive
\[
A \cup B
\] 

intersection - A and B
\[
A \cap B
\] 

##### Ex 

Tennis Cont.

d. $\overline{B} =$ brand does not start w/ W
\[
    \overline{B} = \{ HO, HM, PO, PM, SO, SM \}
\]

