---
title: "Chapter 2"
author: "Logan | Nick"
output: github_document
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.align="center", echo = FALSE)
```

```{r message=FALSE}
if (!require(mosaic)) install.packages("mosaic", repos = 'https://cloud.r-project.org')
library(mosaic) # load the package mosaic to use its functions
```

# 2.1

##### Def

Experiment : any activity / situation w/ uncertainty about which $x \ge 2$ outcomes are possible 

- coin toss, roll a die, draw a card

##### Def

### Sample Space

- collection of all possible outcomes of a chance experiment 

Notate - s or $\mathcal{U}$

- Toss a coin : heads or tails

##### Def

#### Event : 

- any collection of outcomes from a sample space of a chance experiment

Notate

CAP letters : $A, B, C, \ldots$ 

##### Def

Simple Event : event that consists of \underline{one} outcome

Compound Event : event that consists of \underline{more than one} outcome

##### Ex

Tennis : A tennis shop carries 5 brands of rackets (Head, Prince, Sazenger, Wimbledon, Wilson). Each racket comes in midsize / oversize

a. sample space

*insert diagram here*

b. Let A be the event an oversized racket is piurchased
\[
A = \{HO, PO, SO, WimO, WilO\}
\] 

c. Let B be the event the name brand starts w/ a W
\[
 B = \{ WimM, WimO, WilM, WilO \}
\] 

### Forming New Sets

 Let A and B be any 2 events

##### Def

 Complement of A : 

 - all outcomes in S, not in A

##### Notate

$A^{\prime}, \overline{A}, A^c$

##### Notate

union - A or B - inclusive
\[
A \cup B
\] 

intersection - A and B
\[
A \cap B
\] 

##### Ex 

Tennis Cont.

d. $\overline{B} =$ brand does not start w/ W
\[
    \overline{B} = \{ HO, HM, PO, PM, SO, SM \}
\]

e. Head, Prince, and Wilson are US companies. Let C define rackets from the U.S.
\begin{equation*}
\begin{split}
    C &= \{ HO, HM, PO, PM, WilO, WilM \} \\
    B \cup C &=  \{ HO, HM, PO, PM, WilO, WilM, WimO, WimM \}
\end{split}
\end{equation*}

f. List outcomes in $B \cap C$.
\[
B \cap C = \{ WilO, WilM \}
\] 
g. $\overline{(B \cap C)} = \{ HO, HM, PO, PM, WimO, WimM \}$


### Two Mutually Exclusive Events

##### Def

mutually exclusive : no outcomes in common


##### Def

Disjoint : no outcomes in common

*include figure here*

##### Note

If A and B are disjoint, $A \cap B = \emptyset$

*include figure here*

##### Ex

\begin{equation*}
\begin{split}
    A &= \{ 4, 6, 8, 10, 12 \} \quad B = \{ 8, 10, 12, 14 \} \quad C = \{ 12, 14, 16 \} \quad D = \{ 16, 18 \} \\
    A \cap B &= \{ 8, 10, 12 \} \\
    B \cap C &= \{ 12, 14 \} \\
    A \cap (C \cap D) &= A \cap \{ 16 \} = \emptyset \\
    A \cap C &= \{ 12 \} \\
    B \cap D &= \{ \} = \emptyset \\
    (A \cap B) \cup C &= \{ 8, 10, 12 \} \cup C = \{ 8, 10, 12, 14, 16 \} \\
    (A \cap B) \cup (B \cap C) &= \{ 8, 10, 12 \} \cup \{ 12, 14 \} = \{ 8, 10, 12, 14 \}
\end{split}
\end{equation*}

# 2.2 Classical Probability

- N equal likely outcomes

- each outcome has probability $\frac{1}{N}$
