---
title: "Chapter 2"
author: "Logan | Nick"
output: pdf_document
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.align="center", echo = FALSE)
```

```{r message=FALSE}
if (!require(mosaic)) install.packages("mosaic", repos = 'https://cloud.r-project.org')
library(mosaic) # load the package mosaic to use its functions
```

# 2.1

##### Def

Experiment : any activity / situation w/ uncertainty about which $x \ge 2$ outcomes are possible 

- coin toss, roll a die, draw a card

##### Def

### Sample Space

- collection of all possible outcomes of a chance experiment 

Notate - s or $\mathcal{U}$

- Toss a coin : heads or tails

##### Def

#### Event : 

- any collection of outcomes from a sample space of a chance experiment

Notate

CAP letters : $A, B, C, \ldots$ 

##### Def

Simple Event : event that consists of \underline{one} outcome

Compound Event : event that consists of \underline{more than one} outcome

##### Ex

Tennis : A tennis shop carries 5 brands of rackets (Head, Prince, Sazenger, Wimbledon, Wilson). Each racket comes in midsize / oversize

a. sample space

*insert diagram here*

b. Let A be the event an oversized racket is piurchased
\begin{equation*}
\begin{split}
    A = \{HO, PO, SO, WimO, WilO\}
\end{split}
\end{equation*}

c. Let B be the event the name brand starts w/ a W

\begin{equation*}
\begin{split}
    B = \{ WimM, WimO, WilM, WilO \}
\end{split}
\end{equation*}

### Forming New Sets

 Let A and B be any 2 events

##### Def

 Complement of A : 

 - all outcomes in S, not in A

##### Notate

$A^{\prime}, \overline{A}, A^c$

##### Notate

union - A or B - inclusive

\begin{equation*}
\begin{split}
    A \cup B
\end{split}
\end{equation*}

intersection - A and B
\begin{equation*}
\begin{split}
    A \cap B
\end{split}
\end{equation*}

##### Ex 

Tennis Cont.

d. $\overline{B} =$ brand does not start w/ W
\begin{equation*}
\begin{split}
    \overline{B} = \{ HO, HM, PO, PM, SO, SM \}
\end{split}
\end{equation*}

e. Head, Prince, and Wilson are US companies. Let C define rackets from the U.S.
\begin{equation*}
\begin{split}
    C &= \{ HO, HM, PO, PM, WilO, WilM \} \\
    B \cup C &=  \{ HO, HM, PO, PM, WilO, WilM, WimO, WimM \}
\end{split}
\end{equation*}

f. List outcomes in $B \cap C$.

\begin{equation*}
\begin{split}
    B \cap C = \{ WilO, WilM \}
\end{split}
\end{equation*}

g. $\overline{(B \cap C)} = \{ HO, HM, PO, PM, WimO, WimM \}$


### Two Mutually Exclusive Events

##### Def

mutually exclusive : no outcomes in common


##### Def

Disjoint : no outcomes in common

*include figure here*

##### Note

If A and B are disjoint, $A \cap B = \emptyset$

*include figure here*

##### Ex

\begin{equation*}
\begin{split}
    A &= \{ 4, 6, 8, 10, 12 \} \quad B = \{ 8, 10, 12, 14 \} \quad C = \{ 12, 14, 16 \} \quad D = \{ 16, 18 \} \\
    A \cap B &= \{ 8, 10, 12 \} \\
    B \cap C &= \{ 12, 14 \} \\
    A \cap (C \cap D) &= A \cap \{ 16 \} = \emptyset \\
    A \cap C &= \{ 12 \} \\
    B \cap D &= \{ \} = \emptyset \\
    (A \cap B) \cup C &= \{ 8, 10, 12 \} \cup C = \{ 8, 10, 12, 14, 16 \} \\
    (A \cap B) \cup (B \cap C) &= \{ 8, 10, 12 \} \cup \{ 12, 14 \} = \{ 8, 10, 12, 14 \}
\end{split}
\end{equation*}

# 2.2 Classical Probability

- N equal likely outcomes

- each outcome has probability $\frac{1}{N}$.

##### Notate

$P(E)$ is the probability of event $E$ 

\begin{equation*}
\begin{split}
    P(E) = \frac{\text{\# of distinct outcomes in E}}{\text{\# of outcomes in sample space}}
\end{split}
\end{equation*}

### Empirical Probability

- conduct a chance experiment and count the \# of times E occurs

- N is the \# of times the experiment was conducted

\underline{Estimate} of $P(E) = \dfrac{ \text{\# of times E occurs} }{N}$ 

## Law of Large Number's : 

As the number of repetitions of an experiment increases, the chance that the relative frequency of occurrences of an event will differ from the true probability of an event approaches 0.

- Problem : How large is large enough?

##### Note : 

long run stabilization will also occur

*insert graph here*

## Subjective Probability

- a personal measure / belief

  - can be based on evaluation of facts & personal experience

##### Ex

At a hospital, there were $645$ boys born and $721$ girls born

a. Find : experimental probability of having a girl

\begin{equation*}
\begin{split}
    P(girl) = \frac{721}{645 + 721} \approx .53 = 53 \%
\end{split}
\end{equation*}

b. John is @ a cookout and wants to get a drink from the cooler. The cooler contains 12 sodas, 10 waters, & 5 beers.
- Find P(water)
\begin{equation*}
\begin{split}
    P(water) = \frac{10}{12 + 10 + 5} \approx .37 = 37 \% 
\end{split}
\end{equation*}

c. 2 dive rolls : what is $P(second roll > 1^{st} roll)$ 

- using drawing of all probabilities $P(2^{nd} > 1^{st}) = 15 / 36 \approx .42 = 42 \%$

## Basic Properties

1. For any event, the probability it will occur is b/n 0 and 1

2. If $S(Sample space)$, then $P(S) = 1$ 

3. $P(\emptyset) = 0$ 

4. If 2 events $A$ & $B$ are \underline{mutually exclusive}, then $P(A \cup B) = P(A) + P(B)$

5. For any event, $E$, the $P(E) + P( \overline{E} ) = 1$
   - $P( \overline{E} ) = 1 - P(E)$ & $P(E) = 1 - P(\overline{E})$

##### Ex

1. $P( \text{ roll 2 die \& not get doubles }) = 1 - P( \text{ roll doubles } ) = 1 - \frac{6}{36} = 5/6 \approx .83$

2. Find the probability of drawing a Jack or a diamond from a standard 52 card deck

\begin{equation*}
\begin{split}
    P(J \cup \diamondsuit) = P( \diamondsuit ) + P(J) - P(J \cap \diamondsuit) = \frac{13 + 4 - 1}{52} = \frac{16}{52}
\end{split}
\end{equation*}

##### Note
\begin{equation*}
\begin{split}
    P(J \cup \diamondsuit) \ne P( \diamondsuit ) + P(J)
\end{split}
\end{equation*}

3. When rolling 2 die, what is prob. of rolling @ most 1 even number?

\begin{equation*}
\begin{split}
    P( \text{@ most one even} ) = 1 - P( \text{both even} ) = 1 - 9/36 = .75
\end{split}
\end{equation*}

## Union

If $A_1, A_2, \ldots, A_n$ is a collection of disjoint (mutually exclusive) events, then 

\begin{equation*}
\begin{split}
    P(A_1, A_2, \ldots, A_n) = \sum\limits_{i=1}^{\infty} P(A_i) 
\end{split}
\end{equation*}


##### Ex

\begin{table}[h]
    \centering
    \label{tab:label}
    \begin{tabular}{c | c | c | c | c | c | c}
    \# of CD's purchased & 1 & 2 & 3 & 4 & 5 & 6+ \\
    \hline
    prob & .45 & .25 & .1 & .1 & .07 & .03
    \end{tabular}
\end{table}

a. $P(CD \le 3) = P(1) + P(2) + P(3) = .45 + .25 + .1 = .8$ 

b. Prob the next purchase is at most 3 CD's $= P(\# \le 3) P(1) + P(2) + P(3) = .45 + .25 + .1 = .8$

c. What is the prob the next customer buys 5 or more $= P(\# \ge 5) = .07 + .03 = .1$

d. What is $P(\# \ge 2)$ and what does it represent?
\begin{equation*}
\begin{split}
    P(\# \ge 2) = 1 - P(1) = 1 - .45 = .55,
\end{split}
\end{equation*}
which is the event the next customer buys 2 or more CD's

##### Def

General Addition Rule
\begin{equation*}
\begin{split}
    P(A \cup B) &= P(A) + P(B) - P(A \cap B) \\
    P(A \cup B \cup C) &= P(A) + P(B) + P(C) - P(A \cap B) - P(B \cap C) - P(A \cap C) + P(A \cap B \cap C)
\end{split}
\end{equation*}

*insert graph here*

# 2.3

For a multistage experiment w/ $n$ stages where the first stage has $k_1$ outcomes, the second $k_2$ outcomes, $\ldots$, the total number of possible outcomes for the sequence is the multiplication rule.

## Multiplication Rule

\begin{equation*}
\begin{split}
    k_1 \cdot k_2 \cdot \ldots \cdot k_n
\end{split}
\end{equation*}

##### Ex

Determine \# of 5 digit zip codes (digits repeat, first can't be 0)

\begin{equation*}
\begin{split}
    \frac{9 \text{ choices (1 - 9) }}{1^{st} \text{ digit}} &\cdot \frac{10 \text{ (0 - 9) }}{2^{nd} \text{ digit}} \cdot \frac{10 \text{ (0 - 9) }}{3^{rd} \text{ digit}} \cdot \frac{10 \text{ (0 - 9) }}{4^{th} \text{ digit}} \cdot \frac{10 \text{ (0 - 9) }}{5^{th} \text{ digit}} \\
    9 &\cdot 10 \cdot 10 \cdot 10 \cdot 10
\end{split}
\end{equation*}

##### Ex

odd six digit pins if digits can't be repeated

\begin{equation*}
\begin{split}
    5 \cdot 6 \cdot 7 \cdot 8 \cdot 9 \cdot 5 \Leftarrow \text{5th digit must be odd}
\end{split}
\end{equation*}


