---
title: "Chapter 4"
author: Logan | Nick
output: word_document
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.align="center", echo = FALSE, message=FALSE, results = FALSE)
```

```{r message=FALSE}
if (!require(mosaic)) install.packages("mosaic", repos = 'https://cloud.r-project.org')

library(mosaic) # load the package mosaic to use its functions
```

# 4.1 Probabiility Density Functions

##### Def :

A continuous RV is one in which each of the following is true

  * the set of all possible values contains infinite number of possibilities on an interval of the real number line
  
  * no value has positive probability for all x


\[
P(a \le X \le b) = P(a < X \le b) = P(a \le X < b)
\]

  * The pdf can not be listed as a table due to the "number" of outcomes 
  
The pdf of a continuous random variable X given by f(x) and has the following properties

1. $f(x) \ge 0$ for all $x$

2. $\int_{-\infty}^{\infty} f(x)dx = 1$

3. $P(a < X < b) = \int_{a}^{b} f(x)dx$

  * The graph of f(x) lies on or above x-axis
  
```{r}
a = 1
b = 2
xpnorm(q = c(a,b), mean = 0, sd = 1)
  
```

