---
title: "Chapter 1"
author: "Logan | Nick"
output: github_document
fontsize: 12pt
header-includes:
   - \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.align="center", echo = FALSE)
```

```{r message=FALSE}
if (!require(mosaic)) install.packages("mosaic", repos = 'https://cloud.r-project.org')
library(mosaic) # load the package mosaic to use its functions
```

```{r}
library(float)
```


#5.3, 5.4 Estimation

#### Def 

Parameter : a numerical value that describes the population. This value is \underline{fixed but often unknown in practice}.

#### Def

Statistic : a numerical value that describes a sample

- known from sample data

- varies from sample to sample

- use it to estimate the parameter

- use cap letters to denote R.V.

  - $\overline{X}$ is used for the sample mean R.V.

- use lower case letters to denote a calculated value of the statistic 

  - $\overline{x}$ is used for a specific value of the sample mean

#### Def 

The R.V.'s $X_1, X_2, \ldots, X_n$ are said to form a simple random sample of size $n$. If 

1. The $X_i$'s are independent R.V's

2. Every $X_i$ has the same probability distribution

   - independent and identically distributed (iid.)

The idea of a sampling distribution

1. a random sample is selected from a population

2. A statistic is calculated from the random data

   - a statistic is random as it arises from sample to sample according to random selection

   - the distribution of the statistic is called \underline{sampling distribution}

   - the sampling distribution describes all possible values of the statistic and the probability (likelihood) of those values

#### Ex

Let X be the \# of packages being mailed by a random customer.
\begin{table}[h!]
    \centering
    \begin{tabular}{c | c | c | c | c}
    $x$ & 1 & 2 & 3 & 4 \\
    \hline
    $p(x)$ & .4 & .3 & .2 & .1
    \end{tabular}
\end{table}

a. Consider a random sample of size $n = 2$. Let $\overline{X}$ be the \# of packages shipped.

Obtain the pdf of $\overline{X}$

\begin{table}[H]
    \centering
    \begin{tabular}{c | c | c | c}
    $x_1$ & $x_2$ & $p(x_1, x_2)$ & $\overline{x}$ \\
    \hline
    1 & 1 & $.4 * .4 = .16$ & $\cfrac{1 + 1}{2} = 1$ \\[1.7ex]
    1 & 2 & $.4 * .3 = .12$ & $\cfrac{1 + 2}{2} = 1.5$ \\[1.7ex]
    1 & 3 & $.4 * .2 = .08$ & $2$ \\[1.2ex]
    1 & 4 & $.4 * .1 = .04$ & $2.5$ \\[1.2ex]
    \hline
    2 & 1 & .12 & 1.5 \\[0.5ex]
    2 & 2 & .09 & 2 \\[0.5ex]
    2 & 3 & .06 & 2.5 \\[0.5ex]
    2 & 4 & .03 & 3 \\[0.5ex]
    \hline
    \vdots & \vdots & \vdots & \vdots 
    \end{tabular}
\end{table}
\begin{table}[H]
    \centering
    \begin{tabular}{c | c | c | c | c | c | c | c}
    $\overline{x}$ & 1 & 1.5 & 2 & 2.5 & 3 & 3.5 & 4 \\
    \hline
    $p(\overline{x})$ & .16 & .24 & .25 & .2 & .1 & .04 & .01 \\
    \end{tabular}
\end{table}

b. $P( \overline{x} \le 2.5) = .16 + .24 + .25 + .2 = .85$

c. Find $E(\overline{X} = \sum \overline{x}p(\overline{x}) = 2$

#### Note

$E(X) = 2$ 

### Properties of the Sample Mean

Let $X_1, \ldots, X_n$ be simple random sample from a population with mean $\mu$ and standard deviation $\sigma$.

* $E(\overline{x}) = \mu_{\overline{x}} = \ldots = \mu$

* The variance of $\overline{X}$ is 
\begin{equation*}
    \sigma^2 / n = Var(\overline{X})
\end{equation*}


*insert proof here*

So the standard deviation of $\overline{X}$ is 

\begin{equation*}
    \sigma_{\overline{X}} = \frac{\sigma_{x}}{\sqrt{n}} = \frac{\sigma}{\sqrt{n}}
\end{equation*}

